---
title: "model"
author: "Ziming Huang"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(ggplot2)
library(VGAM)
library(lmtest)
library(dplyr)
```

```{r load data}
## Set pwd to the project folder
setwd("~/Desktop/BIOS651/651final/Associations with Clinical Dementia Rating")
## Project choice 2
dementia = read.csv("./dementia/mds_center9661.csv", stringsAsFactors = T)
attributes(dementia)$names
dim(dementia)
```

## Data Cleaning and Variable Selection
```{r}
var <- c("EDUC", "SEX", "MARISTAT", "RACE", "RESIDENC", "DEP", "FEVALAGE", "CDR")
df <- dementia[, var]
## level: 0:Female, 1:Male

## Relevel marriage status: 0:Never married 1:Divorced 2:Married 3:Widowed 4:Separated
df$MARISTAT <-factor(df$MARISTAT, levels = c('Never married','Divorced','Married','Widowed','Separated'))
df$MARISTAT <- relevel(df$MARISTAT, ref = 'Never married')

## Education: I treat it as continous variable here

## Education: Categorical
## based on quantile
quantile(df$EDUC,probs = c(0,0.25,0.5,0.75,1))
cut_p <- c(0, 12, 14, 18, 25)
df$EDUC_cat <- cut(df$EDUC, breaks = cut_p, labels = c('A','B','C','D'))
## Depression: 0:No, 1:Yes

## Race: 0:Non-white 1:White

## CDR:
df$CDR <- factor(df$CDR, levels = c(0.0,0.5,1.0), ordered = F)
df$CDR_ord <- factor(df$CDR, ordered = T)
```

'other' in residenc is an issue
```{r}
t1 = data.frame(table(dementia$RESIDENC,dementia$CDR))
colnames(t1)[1:2] = c('Res','CDR')
(x1 = xtabs(data = t1, Freq ~ Res + CDR))
## there is an association between the choice of res. and CDR level
## no information about other, can't do any inference. So delete it?
x1 = x1[-2,]
(ct <- chisq.test(x1, simulate.p.value = TRUE))
ct$residuals
ct$p.value ## not independent
```

```{r}
df$RESIDENC <- factor(df$RESIDENC, levels = c('Private residence', 'Retirement community', 'Other', 'Assisted living/boarding home/adult family home', 'Skilled nursing facility/nursing home'))
df$RESIDENC <- relevel(df$RESIDENC, ref = 'Other')
```

I didn't check balance here! There are huge unbalance issues within variables such as RESIDENC. 
We can't simply carry out testing and ignore the poor inference power.

## Model fitting:

This is continuous education
```{r}
multi_glm <- vglm(df$CDR ~ df$EDUC, family = multinomial(refLevel = 1))
multi_glm_null <- vglm(df$CDR ~ 1, family = multinomial())
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

categorical education
```{r}
multi_glm <- vglm(df$CDR ~ df$EDUC_cat, family = multinomial(refLevel = 1))
multi_glm_null <- vglm(df$CDR ~ 1, family = multinomial())
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

Doesn't really matter which one we use. 

```{r}
multi_glm <- vglm(df$CDR_ord ~ df$EDUC_cat, family = cumulative(link = 'logitlink', parallel = T))
multi_glm_null <- vglm(df$CDR_ord ~ 1, family = cumulative(link = 'logitlink', parallel = T))
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

```{r}
multi_glm <- vglm(df$CDR_ord ~ df$EDUC, family = cumulative(link = 'logitlink', parallel = T))
multi_glm_null <- vglm(df$CDR_ord ~ 1, family = cumulative(link = 'logitlink', parallel = T))
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

Cummulative works much better! 
We can double check the ordinal trend by gamma test or Choctron-Armitage

I will just use educ as categorical here

```{r}
fit0 <- vglm(data = df, 
             CDR_ord ~ EDUC_cat +SEX+ MARISTAT + RACE + RESIDENC + DEP + FEVALAGE,
             family = cumulative('logitlink', parallel = T))
summary(fit0)
```

we need to account for the proportionality assumption !!!

Based on the main model we might want to go back to literature review and modify the form of variables of interests. Overall it looks pretty good.

Recode marriage and residence type
```{r}
df$RES_PRIVATE <- df$RESIDENC == 'Private residence'
df$Marriage <- df$MARISTAT == 'Married'
```

for main model:
We might want to add a few more without education according to lit. review
1. need gof: aic/bic/deviance/hltest
2. need interpretation!!! 

```{r}
main_model <- vglm(data = df, 
             CDR_ord ~ EDUC_cat +SEX+ Marriage + RACE + RES_PRIVATE + DEP + FEVALAGE,
             family = cumulative('logitlink', parallel = T))
summary(main_model)
```
still need gof test (HLTEST)!

We might want to think about race again
```{r}
table(df$RACE)
```
The unblance here is quite significant. rid of race can also makes our lives easier.


## Interactions: how you guys want to do this?
```{r}
fit0 <- vglm(data = df, 
             CDR_ord ~ EDUC_cat*(SEX+Marriage+ DEP + FEVALAGE + RES_PRIVATE + RACE),
             family = cumulative('logitlink', parallel = T))
summary(fit0)
```

```{r}
step4vglm(fit0, direction = 'backward')
```

According to backward selection + literature review: 
CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + EDUC_cat:RES_PRIVATE + Race
is the best one we get.

We might want to add a few more without education according to lit. review
1. need gof: aic/bic/deviance/hltest (a lot of work):
2. need model diagnostic: residual,dfbetas,dffit,cooks dist... (visualizaion):
      unbalance everywhere, proportional assumption not hold, 
      (link function might not be logit cloglog or loglog)
3. need interpretation!!! (a lot of work):


```{r}
summary(vglm(formula = CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + RACE +EDUC_cat:RES_PRIVATE, family = cumulative("logitlink", 
    parallel = T), data = df))
```

```{r}
library(MASS)
MF <- polr(CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + RACE+  FEVALAGE + 
    RES_PRIVATE + EDUC_cat:RES_PRIVATE, data = df, method = 'probit')
MF
```

```{r}
OIM <- polr(CDR_ord ~ 1, data = df)
anova(OIM,MF)
```

```{r}
sum(predict(MF)==df$CDR)
```


## inference is bad at 0.5 group
```{r}
sum(predict(MF) == df$CDR)
```

```{r}
df$s1 <- (as.numeric(df$CDR)<=1)*1
df$s2 <- (as.numeric(df$CDR)<=2)*1
```

```{r}
sm1 <- glm(data = df, s1~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + RACE +EDUC_cat:RES_PRIVATE,family = binomial())
sm2 <- glm(data = df, s2~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + RACE +EDUC_cat:RES_PRIVATE,family = binomial())
pm1 <- predict(sm1)
a=exp(pm1)/(1+exp(pm1))
pm2 <- predict(sm2)
b=exp(pm2)/(1+exp(pm2))
d <- cbind(a,b-a,1-b)
e = apply(d, 1, which.max)
sum(df$CDR == (e-1)*0.5)
```

```{r}
chisq.test(df$CDR,predict(MF))
```

```{r}
AIC(MF)
BIC(MF)
PseudoR2(MF, which = c("CoxSnell","Nagelkerke","McFadden"))
```

















