---
title: "model"
author: "Ziming Huang"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(ggplot2)
library(VGAM)
library(lmtest)
library(dplyr)
library(generalhoslem)
library(gofcat)
library(DescTools)
```

```{r load data}
## Set pwd to the project folder
setwd("~/Desktop/BIOS651/651final/Associations with Clinical Dementia Rating")
## Project choice 2
dementia = read.csv("./dementia/mds_center9661.csv", stringsAsFactors = T)
dim(dementia)
```

## Data Cleaning and Variable Selection
```{r}
var <- c("EDUC", "SEX", "MARISTAT", "RACE", "RESIDENC", "DEP", "FEVALAGE", "CDR", "RLDEM")
df <- dementia[, var]

## Relevel marriage status: 0:Never married 1:Divorced 2:Married 3:Widowed 4:Separated
df$MARISTAT <-factor(df$MARISTAT, levels = c('Never married','Divorced','Married','Widowed','Separated'))
df$MARISTAT <- relevel(df$MARISTAT, ref = 'Never married')

## Residence type
df$RESIDENC <- factor(df$RESIDENC, levels = c('Private residence', 'Retirement community', 'Other', 'Assisted living/boarding home/adult family home', 'Skilled nursing facility/nursing home'))
df$RESIDENC <- relevel(df$RESIDENC, ref = 'Other')

## Education coded as categorical variable based on the quantile
quantile(df$EDUC,probs = c(0,0.25,0.5,0.75,1))
## adjust it to give it meaning
cut_p <- c(0, 12, 14, 18, 25)
df$EDUC_cat <- cut(df$EDUC, breaks = cut_p, labels = c('A','B','C','D'))
## CDR:
df$CDR <- factor(df$CDR, levels = c(0.0,0.5,1.0), ordered = F)
df$CDR_ord <- factor(df$CDR, ordered = T)
```

## analysis on residenc
```{r}
t1 = data.frame(table(dementia$RESIDENC,dementia$CDR))
colnames(t1)[1:2] = c('Res','CDR')
(x1 = xtabs(data = t1, Freq ~ Res + CDR))
x1 = x1[-2,]
(ct <- chisq.test(x1, simulate.p.value = TRUE))
ct$residuals
ct$p.value ## not independent

## To deal with the unbalance and other, we combine res into private vs non-private
df$RES_PRIVATE <- df$RESIDENC == 'Private residence'
t2 = data.frame(table(df$RES_PRIVATE,dementia$CDR))
colnames(t2)[1:2] = c('Res','CDR')
(x2 = xtabs(data = t2, Freq ~ Res + CDR))
x2 = x2[-2,]
(ct <- chisq.test(x2, simulate.p.value = TRUE))
ct$residuals
ct$p.value ## not independent, we still preserve the association. Binary res seems valid!
```


```{r}
table(df$MARISTAT)
t3 = data.frame(table(dementia$MARISTAT,dementia$CDR))
colnames(t3)[1:2] = c('marriage','CDR')
(x3 = xtabs(data = t3, Freq ~ marriage + CDR))
(ct <- chisq.test(x3, simulate.p.value = TRUE))
ct$residuals
ct$p.value ## not independent
## To address this we combine marriage into binary 

df$Marriage <- df$MARISTAT == 'Married'
t4 = data.frame(table(df$Marriage,dementia$CDR))
colnames(t4)[1:2] = c('marriage','CDR')
(x4 = xtabs(data = t4, Freq ~ marriage + CDR))
(ct <- chisq.test(x4, simulate.p.value = TRUE))
ct$residuals
ct$p.value ## not independent

```


## Model fitting:

This is continuous education
```{r}
multi_glm <- vglm(df$CDR ~ df$EDUC, family = multinomial(refLevel = 1))
multi_glm_null <- vglm(df$CDR ~ 1, family = multinomial())
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

categorical education
```{r}
multi_glm <- vglm(df$CDR ~ df$EDUC_cat, family = multinomial(refLevel = 1))
multi_glm_null <- vglm(df$CDR ~ 1, family = multinomial())
waldtest(multi_glm, multi_glm_null)
lrtest_vglm(multi_glm, multi_glm_null)
```

Doesn't really matter which one we use. Our main covariate education is significant! We want to use
categorical to increase intepretability.

## Cummulative model:
There is obvious trend within cdr, so we want to check the cummulative model for it!

```{r}
cumu_glm <- vglm(df$CDR_ord ~ df$EDUC_cat, family = cumulative(link = 'logitlink', parallel = F))
cumu_glm_null <- vglm(df$CDR_ord ~ 1, family = cumulative(link = 'logitlink', parallel = F))
waldtest(cumu_glm, cumu_glm_null)
lrtest_vglm(cumu_glm, cumu_glm_null)
```

further we want to check if the effect of education is homogenous
```{r}
cumu_glm_T <- vglm(df$CDR_ord ~ df$EDUC_cat, family = cumulative(link = 'logitlink', parallel = T))
lrtest_vglm(cumu_glm, cumu_glm_T)
```
No, it is not statistically correct to assume proportionality assumption.

## Main model regression:

we first want to check vif before adding confounders (based on literature review)
```{r}
df$s1 <- (as.numeric(df$CDR)<=1)*1
df$s2 <- (as.numeric(df$CDR)<=2)*1
sm1 <- glm(data = df, s1~EDUC_cat + SEX + MARISTAT + DEP + FEVALAGE + 
    RES_PRIVATE,family = binomial())
sm2 <- glm(data = df, s2~EDUC_cat + SEX + MARISTAT + DEP + FEVALAGE + 
    RES_PRIVATE + RACE,family = binomial())
```
There is no colinearity 
```{r}
main_model_F <- vglm(data = df, 
             CDR_ord ~ EDUC_cat +SEX+ Marriage+ RACE + RES_PRIVATE + DEP + FEVALAGE,
             family = cumulative('logitlink', parallel = F))
summary(main_model_F)
logLik.vlm(main_model_F)
```
The loglik is not reasonable! The non-parallelism assumption has resulted in intersecting linear/additive predictors. 

```{r}
## To avoid the unreasonable model, i wrote it myself
cumf_m1 <- glm(data = df, s1~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE, family = binomial())
cumf_m2 <- glm(data = df, s2~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + RACE,family = binomial())
ll1 <- logLik(cumf_m1)
ll2 <- logLik(cumf_m2)
ll1+ll2
```

```{r}
main_model_T <- vglm(data = df, 
             CDR_ord ~ EDUC_cat +SEX+ Marriage+ RACE + RES_PRIVATE + DEP + FEVALAGE,
             family = cumulative('logitlink', parallel = T))
logLik.vlm(main_model_T)
```

we will stick to this model = T

```{r}
main_model <- vglm(data = df, 
             CDR_ord ~ EDUC_cat +SEX+ Marriage + RACE + RES_PRIVATE + DEP + FEVALAGE,
             family = cumulative('logitlink', parallel = T))
summary(main_model)
```
## Interactions: 
Add interaction based on aic
```{r}
fit0 <- vglm(data = df, 
             CDR_ord ~ EDUC_cat*(SEX+Marriage+ DEP + FEVALAGE + RES_PRIVATE + RACE),
             family = cumulative('logitlink', parallel = T))
summary(fit0)
```

```{r}
step4vglm(fit0)
```



According to both direction selection + literature review: 
CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + FEVALAGE + RACE
    RES_PRIVATE + EDUC_cat:FEVALAGE + EDUC_cat:RES_PRIVATE
is the best one we get.


```{r}
fm <- vglm(CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + FEVALAGE + RACE+
    RES_PRIVATE + EDUC_cat:FEVALAGE + EDUC_cat:RES_PRIVATE, data = df,family = cumulative("logitlink", 
    parallel = T))
summary(fm)
```

```{r}
## check linearity for age
df$FEVALAGE2=df$FEVALAGE^2
fm2 <- vglm(CDR_ord ~ EDUC_cat + SEX + Marriage + DEP + FEVALAGE + FEVALAGE2 + RACE+
    RES_PRIVATE + EDUC_cat:FEVALAGE + EDUC_cat:RES_PRIVATE, data = df,family = cumulative("logitlink", 
    parallel = T))
lrtest_vglm(fm2,fm)
```
The non-linear effect is not significant 

```{r}
AIC(fm)
BIC(fm)

## not necessary
n = dim(df)[1]
prob_table <- exp(predict(fm))/(1+exp(predict(fm)))
prob_table <- cbind(prob_table[,1], prob_table[,2]-prob_table[,1], 1-prob_table[,2])
prediction = apply(prob_table, 1, which.max)
(accuracy = sum(df$CDR == (prediction-1)*0.5)/n)
```

```{r}
## accuracy for parrellal = F coding
sm1 <- glm(data = df, s1~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + EDUC_cat:FEVALAGE+RACE +EDUC_cat:RES_PRIVATE,family = binomial())
sm2 <- glm(data = df, s2~EDUC_cat + SEX + Marriage + DEP + FEVALAGE + 
    RES_PRIVATE + RACE +EDUC_cat:FEVALAGE+EDUC_cat:RES_PRIVATE,family = binomial())
pm1 <- predict(sm1)
a=exp(pm1)/(1+exp(pm1))
pm2 <- predict(sm2)
b=exp(pm2)/(1+exp(pm2))
d <- cbind(a,b-a,1-b)
e = apply(d, 1, which.max)
e= apply(d, 1, which.max)
(accuracy = sum(df$CDR == (e-1)*0.5)/n)
```

```{r}
# McFadden's pseudo-R^2
null_fm <- vglm(CDR_ord ~ 1, data = df,family = cumulative("logitlink", 
    parallel = T))
lrtest_vglm(null_fm, fm)
print(pseudo_R2 <- 1 - deviance(fm) / deviance(null_fm))
```

## Hoslem
```{r}
logitgof(df$CDR, fitted(fm), ord= T, g = 5)
```

we pick several g 

variance covariance plot check for colinearity 

```{r}
VIF(sm1)
VIF(sm2)
```



















